---
title:  "Lab 3: Visualize and wrangle data"
subtitle: "Not graded, just practice"
date:   2023-09-14
author: Katie Schuler
number-sections: true
number-depth: 3
---


```{r}
#| echo: false
#| message: false

library(webexercises)
library(tidyverse)
library(languageR)
```


Todays lab will make use of the `ratings` data in the `languageR` package. `glimpse()` of data frame is provided for your reference. 

```{r}
ratings %>% glimpse()
```

## `ggplot2` 


Given code blocks a, b, and c; and the plot below:

```{r}
#| output: false
# CODE BLOCK a ---------------------------#
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity)
    ) +
    geom_point(color = "blue")

# CODE BLOCK b ---------------------------#
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity, color = "blue")
    ) 

# CODE BLOCK c ---------------------------#
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity)
    ) +
    geom_point()
```

```{r}
#| echo: false
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity)
    ) +
    geom_point(color = "blue") +
    labs(tag="A")
```

Which of the code blocks above generate this plot?

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer="a", "b", "c"
)

cat(longmcq(choices))
```

In plot A above, is the color aesthetic mapped, set, or both? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
   "mapped", answer="set", "both"
)

cat(longmcq(choices))
```

In plot A above, which of the following aesthetics should we set to make the points more transparent?


```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
   "color", "fill", answer="alpha", "shape"
)

cat(longmcq(choices))
```

In plot A above, which of the following would change the x axis label to "FQ"?

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
   answer ="add a `labs()` layer with `x='FQ'` argument",
   "add a `labs()` layer with `y='FQ'` argument",
   "change the mapping argument from `x='Frequency'` to `x='FQ'`" 
)

cat(longmcq(choices))
```


```{r}
#| echo: false
#| message: false
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity)
    ) +
    geom_point(color = "red") +
    geom_smooth(method = "lm") +
    labs(tag="B") 
```

In plot B above, which geom(s) are used to represent the data? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
   "geom_histogram()", "geom_density()", "geom_bar()", answer="geom_point()", answer="geom_smooth()"
)

cat(longmcq(choices))
```

True or false, the blue line in plot B above is mapped to the Class aesthetic? 


```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c("True", answer="False")

cat(longmcq(choices))
```

In plot B above, which of the following variables is mapped to the x aesthetic? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer="Frequency", 
    "meanFamiliarity", 
    "FreqSingular",
    "FreqPlural"
)

cat(longmcq(choices))
```

True or false, in plot B above, the default statistical transformation in the geom responsible for the red dots is "identity".

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer="True", "False")

cat(longmcq(choices))
```

Suppse we run the code below. Which of the following plots will be returned? 

```r
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity, color = Class)
    ) +
    geom_point() +
    geom_smooth(method = "lm", color = "red") 
```

```{r}
#| echo: false
#| message: false
#| layout-ncol: 2
#| layout-valign: bottom
#| cap-location: bottom
ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity, color = Class)
    ) +
    geom_point(color = "red") +
    geom_smooth(method = "lm") +
    labs(tag="B") 

ggplot(
    data = ratings, 
    mapping = aes(x = Frequency, y = meanFamiliarity, color = Class)
    ) +
    geom_point() +
    geom_smooth(method = "lm", color = "red") +
    labs(tag="B") 
```



## `dplyr` 


Which of the following `dplyr` functions retuns a data frame? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer ="`select()`", 
    answer ="`mutate()`", 
    answer ="`filter()`",
    answer ="`rename()`",
    "None of the above"
)

cat(longmcq(choices))
```


Which of the following `dplyr` functions takes a number as their first argument? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "`select()`", 
    "`mutate()`", 
    "`filter()`",
    "`rename()`",
    answer = "None of the above"
)

cat(longmcq(choices))
```


True or false, the following code blocks are equivalent. 

```r
# option 1
ratings %>% select(Word, Frequency) %>% glimpse()

# option 2
glimpse(select(ratings, Word, Frequency))

```

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer="True", "False")

cat(longmcq(choices))
```


 

True or false, the following code options are equivalent 


```r
# option 1
ratings %>% 
    select(Word:Class) %>% 
    mutate(Length/Frequency, .after = Class)

# option 2
ratings %>% 
    select(Word:Class) %>% 
    mutate(Length/Frequency)

```

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer="True", "False")

cat(longmcq(choices))
```



Suppose you run the following code block. How many rows would be in the resulting data frame? `r fitb(2, width =4)`

```r 

ratings %>% group_by(Class) %>% summarise(n = n())

```


Given the code block in the previous question, what will `n()` do? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "summarize all classes including the letter n",
    answer = "counts the number of rows per Class",
    "adds the string `n` before each value of Class",
    "error: missing arguments to n()"

)

cat(longmcq(choices))
```

::: {.callout-warning title="Under Construction"}
We will add more practice questions tomorrow so you have more practice to study for Tuesdays Quiz!!
:::





<!-- Which of the following code blocks returned this data frame?  -->


<!-- ### Columns 

Suppose you used mutate to add a column dividing Frequency by FamiliySize and the following data frame was returned. Which of the following arguments were specified? 

Suppose you run the following code. How many of the 14 columns in the original dataset would be returned? 

True or false, the following code blocks are equivalent. 

rename and select rename 

### Group and summarise

True or false, `group_by()` generates a grouped data frame. 




True or false, the following code blocks are equivalent. 
 -->





