---
title:  "Practice quiz 4"
subtitle: "Not graded, just practice"
date:   2023-11-28
author: Katie Schuler
number-sections: false
---

```{r}
#| echo: false
#| message: false

library(webexercises)
library(tidyverse)
library(optimg)
library(tidymodels)
theme_set(theme_gray(base_size = 16))
set.seed(60)

data <- tibble(
    y = rnorm(1000, 5, 1), 
    noise = rnorm(1000, 10, 4), 
    x = y + 10 + noise
)
```

## 1 Model reliability 

a. As we collect more data, our paramter estimates

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer = "become more reliable",
    "become less reliable", 
    "stay the same"
)

cat(longmcq(choices))
```

b. Each figure below plots 100 bootstrapped models. Which figure includes more data points? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "Figure A",
    answer = "Figure B", 
    "Both figures have the same number of data points"
)

cat(longmcq(choices))
```

c. As we collect more data, what happens to the confidence interval around our parameter estimates? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "It gets bigger (wider)",
    "It stays the same", 
    answer = "It gets smaller (narrower)"
)

cat(longmcq(choices))
```

d. True or false, we can obtain confidence intervals around parameter estimates for models in the same we we did for point estimates like the mean. 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer = "True", "False")

cat(longmcq(choices))

```

e. Model reliability asks how certain we can be about our parameter estimates. Why is there uncertainty around our parameter estimates? 

```{r}
#| code-fold: true
#| code-summary: "answer"
#| eval: false
"Because we are interested in the model parameters that 
best describe the population from which the sample was
 drawn. Due to sampling error, we can expect some 
 variability in the model parameters."

```

f. The figure below shows the model fit for a sample of 10 participants. Suppose we repeated our experiment with 10 new participants. True or false, fitting the same model to these new data would yield approximately the same parameter estimates. 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c("True", answer = "False")

cat(longmcq(choices))

```

g. True or false, a model with high accuracy must also have high reliability.

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c("True", answer="False")

cat(longmcq(choices))

```


## Nonlinear models 

a. The model `y ~ poly(x,2)` is plotted in which of the figures below? 





b. Which of the equations below expresses a quadratic polynomial?  

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "A",
    "B",
    answer="C", 
    "D"
)

cat(longmcq(choices))
```

c. True or false, `lm()` can be used to fit a linearized nonlinear model. 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer="True", "False")

cat(longmcq(choices))

```

d. Fill in each blank below with the model building process best described by the definition: 

    - `r mcq(c("model specification", answer =  "model fitting", "model accuracy", "model reliability"))` is finding the best fitting parameter estimates.
    - `r mcq(c("model specification", "model fitting", "model accuracy", answer =  "model reliability"))` is quantifying the uncertainty on the parameter estimates.
    - `r mcq(c(answer =  "model specification", "model fitting", "model accuracy", "model reliability"))` is choosing the type of model and its functional form.
    - `r mcq(c("model specification", "model fitting", answer =  "model accuracy", "model reliability"))` is quantifying how well our model fits our data. 


## 3 Classification 

a. Which of the following aspects of model building apply to classification models? (Choose all that apply)

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer="specify",
    answer="fit",
    answer="accuracy", 
    answer="reliability"
)

cat(longmcq(choices))
```

b. What is the difference between regression and classification? 

```{r}
#| code-fold: true
#| code-summary: "answer"
#| eval: false
"Regression predicts a continuous response varaible, 
classification predicts a discrete response variable"

```

c. Which accuracy metric is best applied to classification models? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "R squared",
    "RMSE",
    answer ="Pecent correct",
    "All of the above"
)

cat(longmcq(choices))
```

d. In the figure below, which aspect shows the response variable? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "x-axis",
    "y-axis",
    answer ="color of points",
    "blue line"
)

cat(longmcq(choices))
```

e. Which figure below could be a classification model? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "Figure A",
    answer ="Figure B",
    "both"
)

cat(longmcq(choices))
```


f. Name two kinds of linear classifiers

```{r}
#| code-fold: true
#| code-summary: "answer"
#| eval: false
"Any 2 of those mentioned in class: 
Logistic regression
Linear discriminant analysis (LDA)
Linear support vector machines (SVM)
Nearest-prototype classifiers
Naive Bayes classifiers
"

```


## Classification in R 

a. We can impliment classification via `r mcq(c("linear regression", "cubic polynomials", answer =  "logistic regression", "nearest-neighbor regression", "all of the above"))`

b. True or false, in R, we can perform logistic regression with a generalized linear model. 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(answer="True", "False")

cat(longmcq(choices))

```

c. What 3 elements do all GLMs have? 

```{r}
#| code-fold: true
#| code-summary: "answer"
#| eval: false
"1. A particular distribution for modeling the response variable
2. A linear model 
3. A link function
"

```

d. What is the link function for logistic regression? 

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    answer = "logistic function", 
    "polynomial expansion",
    "log transformation",
    "inverse transformation"
)

cat(longmcq(choices))

```

f. Which of the following fits a logistic regression model in R? Choose one.

```{r}
#| echo: false
#| results: asis

# Define the answer choices
choices <- c(
    "code A", 
    "code B",
    "code C",
    "code A and B only",
    answer = "all of the above"
)

cat(longmcq(choices))

```












