---
title: "Data visualization 1"
date: 09/07/2023
author: 
    - Katie Schuler

---

::: {.callout-warning title="Under Construction"}
:::

## Acknowledgements 

Today's lectures notes and [slides](/slides/data-visualization.qmd) were influenced by and adapted from [Ch 2 in R for Data Science]() and various materials by Dr. Colin Rundel at Duke

## Why visualize data? 

> "Visualization is a fundamentally human activity. A good visualization will show you things you did not expect or raise new questions about the data. A good visualization might also hint that you’re asking the wrong question or that you need to collect different data." -- R for Data Science





## Data visualization with ggplot2
<!---
 
Prerequisite packages. ggplot part of tidyverse package. Already installed on Google Colab's R; just use library.

```{r}
# load the tidyverse and palmer penguins libraries
library(tidyverse)
library(palmerpenguins)
```


We will also be using palmerpenguins package; show here how to use glimpse() to get a look or head() -- can also show on Tue 


```{r}
glimpse(penguins)
```

```{r}
head(penguins)
```

plot we want to replicate: 

here 

Start with data layer; 

```{r}
ggplot(data = penguins) 

```


sample basic plot: 

```{r}

```
 
--->

three main parts: 

- **data**- your (tidy) data
- **aesthetic mappings (aes)** - make data visible
- **geometric objects** - 

other layers: 

- **facets** 
- **statistics** - statistical transformations of data
- **scales** - map data values to visual values of aestetic
- coordinates
- **themes** - overall visuals

special notes for Google Colab: 
- setting size 



## Aesthetic mappings

Practical example: 
- adding type of penguins 




**aesthetic mappings** (aes), make data visible

- `x`, `y`: variable on x and y axis
- `color`: outline color of geom
- `fill`: fill color of geom
- `group`: group geom belongs to
- `shape`: shape used to plot point (circle, square, etc)
- `linetype`: type of line used (solid, dash, et)
- `size`

## Geometric objects


**geometric objects (geoms)** - type of plot

- show adding geom_smooth 



## Additional layers

We'll cover some in the class: 
- adding labels 
- adding theme
- adding fill 

- **facets**
    - `facet_grid(.~var1)`
    - `facet_grid(var2~.)`
    - `facet_grid(var2~var1)`
    - `facet_wrap(~var1)` - wrap facets
    - "scales"? 
- **coordinates**
    - `coord_cartesian(xlim=c(0,5))` - xlim, ylim
    - `coord_flip()`
- **labels** 
    - `labs(title="plot title")` - x, y, subtitle
    - `annotate()`
- **statistics**
- **scales**
- **position adjustments**
- **themes**
    - `theme(legend.position="bottom")`

## Saving plots 

Helper functions: 

- `last_plot()` - returns the last plot
- `ggsave("plot.png", width=5, height=5)` - saves last plot

### Further reading and references

Useful resources:

- ggplot cheat sheet
- [introduction to palmerpenguins](https://allisonhorst.github.io/palmerpenguins/articles/intro.html) 

Recommended further reading: 

- [Ch 2 Data visualization](https://r4ds.hadley.nz/data-visualize) in R for Data Science
- https://moderndive.com/2-viz.html
- [Layers](https://r4ds.hadley.nz/layers) in R for Data Science


Other references: 

- https://r.qcbs.ca/workshop03/book-en/the-basics-of-visualizing-data.html
- https://r4ds.hadley.nz/layers

