---
title:  "Quiz 1"
subtitle: "Data Science for Studying Language & the Mind"
# author: 
#     name: Katie Schuler
#     affiliation: University of Pennsylvania
number-sections: false
format: 
  html: default
  pdf: default

---

::: {.callout-tip}
## Estimated time: 30 minutes

You may need more time if programming is completely new to you, or less if you have some experience already.
:::

**Instructions**

- The quiz is closed book/note/computer/phone
- If you need to use the restroom, leave your exam and phone with the TA
- You have 60 minutes to complete the quiz. If you finish early, you may turn in your quiz and leave early

```{r}
#| echo: false
#| message: false
library(tidyverse)
```

\newcommand\answerbox{%%
    \framebox(400,50){}}

\begin{flushleft}
\makebox[12cm]{\textbf{Name}:\ \hrulefill}
\medskip
\makebox[12cm]{\textbf{PennKey}:\ \hrulefill}
\medskip
\makebox[12cm]{\textbf{Lab section TA}:\ \hrulefill}
\end{flushleft}

\begin{center}
\textbf{Score by topic area}\\
\begin{tabular}{|l|l|}\hline
R basics: general &   \\ \hline
R basics: vectors, operations, subsetting &  \\ \hline
Data importing &  \\ \hline
Data visualization: basics &  \\ \hline
Data visualization: layers  &  \\ \hline
Data wrangling &  \\ \hline
Total & \\ \hline

\end{tabular}
\end{center}

\clearpage

## 1 R basics: general 

(a) Suppose you run the following code. Which command could you run to remove the `y` variable from the current environment? Choose all that apply.

    ```r
    x <- 1 + 2 
    y <- 3 + 4
    z <- 0
    ```
    - [ ] `ls(y)`
    - [ ] `rm(list = ls())`
    - [ ] `rm(y)` 
    - [ ] `remove(y)`

(b) Write an expression that would assign the value `10` to the variable name `my_var`? 

    \answerbox

(c) Which of the following would install the `praise` package? Choose all that apply.
    - [ ] `library(praise)`
    - [ ] `install(praise)`
    - [ ] `install.packages("praise")`
    - [ ] `install.packages(praise)`

(d) Suppose you run the following code. Which functions would return the structure of the object you defined? Choose one.

    ```r
    x <- c("bus", "stop")
    ```
    - [ ] `length(x)`
    - [ ] `print(x)`
    - [ ] `str(x)`
    - [ ] `attributes(x)` 
    - [ ] None of the above

\clearpage 

## 2 R basics: vectors, operations, subsetting 

(a) Suppose you run the following code. What will length(x) return? Write your answer in the box below.

    ```r
    x <- 1:5
    ```
    \answerbox



(b) Suppose you run the following code. What will `y > 4` return? Write your answer in the box below and **show your work**.

    ```r
    x <- seq(2, 8, by = 2)
    y <- x[c(-4)]
    ```
    \answerbox

(c) Suppose you run the following code. What will `typeof(x)` return? 

    ```r
    x <- c("true", "false", "true", "false")
    ```
    - [ ] `double`
    - [ ] `integer`
    - [ ] `character`
    - [ ] `logical`

(d) Suppose you create the following data frame and assign it to the `df` variable. What will `sum(df$antique)` return?

    ```{r}
    #| echo: false


    df <- data.frame(
        antique = 1:3,
        age = 2:4,
        show = c("a", "b", "c")
        
    )
    print(df)

    ```

    - [ ] NULL
    - [ ] Error: no columns include value "antique"
    - [ ] 6
    - [ ] 9 
    - [ ] 15

## 3 Data importing

(a) Which of the following will load the `readr` package into the current environment? Choose all that apply.
    - [ ] `library(tidyverse)`
    - [ ] `library(readr)`
    - [ ] `install.packages("tidyverse")`
    - [ ] `install.packages("readr")`
    - [ ] `import(tidyverse)`
    - [ ] `import(readr)`

(b) Suppose `print(x)` returns the following. What will `is.data.frame(x)` return? Write your answer in the box below.

    ```{r}
    #| echo: false


    x <- tibble(
        x = 1:4,
        y = 5:8,
        z = 9:12
    )
    print(x)
    ```

    \answerbox

(c) Suppose you import "junesales.csv", shown below, with the following code. What would `data$Sale` return? Choose one.

    ```
    Year, Month, Day, Sale
    2023, June, 1, 0
    2023, June, 2, 1
    2023, June, 3, 0
    2023, June, 4, 1
    ```

    ```r
    data <- read_csv("junesales.csv",
        col_types = list(Sale = col_logical())
    )
    ```
    - [ ] A `double` vector with values `0 1 0 1`
    - [ ] A `logical` vector with values `FALSE TRUE FALSE TRUE`
    - [ ] A double vector with values `NA NA NA NA`
    - [ ] `NULL`

(d) Suppose you import a dataset with `readr`, but when you `print(data)` you notice that the `age` column was identified as `character` when you were expecting `double`. Given the resulting tibble, which of the following arguments could you include in blank in the code below to solve this problem?


    ```{r}
    #| echo: false
    data <- tibble(
        age = c(18, "na", 25, 21), 
        graduated = c(F, F, T, T), 
        gpa = c(NA, 3.8, 2.9, 3.1)
    )
    print(data)
    ```

    ```r
    data <- read_csv("data.csv", ________)
    ```

    - [ ] `.drop = NA`
    - [ ] `skip = 1`
    - [ ] `guess_max = Inf`
    - [ ] `na = c("na")`
    - [ ] `col_names = FALSE`

\clearpage

## 4 Data visualization: basics 

Section 5 makes use of the `durationsGe` dataset and plots A, B, and C in the appendix.


(a) Which of the plots above (A and B) did the code blocks below generate?

    ```r
    # code 1
    ggplot(durationsGe, aes(x = DurationOfPrefix, fill = Sex)) +
        geom_density(fill = "lightgray") +
        theme_classic(base_size = 12) +
        labs(y = "") +
        scale_fill_manual(values = c("white", "gray", "black"))
    # code 2
    ggplot(durationsGe, aes(x = DurationOfPrefix)) +
        geom_density(fill = "lightgray") +
        theme_classic(base_size = 12) +
        labs(y = "")
    ```
    - [ ] Code 1 generates plot A, code 2 generates plot B
    - [ ] Code 2 generates plot A, code 1 generates plot B
    - [ ] Code 1 and 2 both generate plot A
    - [ ] Code 1 and 2 both generate plot B 
    - [ ] Code 1 and 2 generate neither plot A nor plot B

(b) Which geoms could be depicted in plots A and B? Choose all that apply. 

    - [ ] `geom_histogram()`
    - [ ] `geom_smooth()`
    - [ ] `geom_line()`
    - [ ] `geom_density()`
    - [ ] `geom_bar()`

\clearpage

(c) True or false, the following code blocks generate the same figure. Write your answer in the following box and **explain why**.

    ```r
    # code block 1
    ggplot(
        data=durationsGe, 
        mapping = aes(y = DurationOfPrefix, x = Sex)) + 
            geom_bar(stat = "identity") 

    # code block 2
      ggplot(
        aes(y = DurationOfPrefix, x = Sex),
        durationsGe) + 
            geom_bar(stat = "identity") 

    ```

    \answerbox


(d) The code below makes use of a new geom, `geom_rug()`, to generate plot C, in which each individual data point is plotted along the x-axis like a "rug". In the box below, rewrite the code such the color of the rug is mapped to the `Sex` variable and the bars of the histogram are filled in with the color "lightblue". 

    ```r
    ggplot(durationsGe, aes(x = DurationOfPrefix)) +
        geom_rug() + 
        geom_histogram() +
        theme_classic(base_size = 12) 

    ```

    \answerbox


\clearpage

## 5 Data visualization: layers

Section 5 makes use of the `durationsGe` dataset and plots D, E, and F in the appendix.



(a) Which of the following would add a small amount of random noise around each point in plot D? Choose all that apply.

    - [ ] add the argument `position = "jitter"` to `geom_point()`
    - [ ] add the argument `position = "random"` to `geom_point()`
    - [ ] add the argument `rand_noise = TRUE` to `geom_point()`
    - [ ] replace `geom_point()` with `geom_jitter()`
    - [ ] replace `geom_point()` with `geom_noise()` 


(b) Which of the following could change plot D to plot E? Choose all that apply

    - [ ] add `facet_wrap(~Sex)`
    - [ ] add `facet_wrap(~Sex, ncol = 2)`
    - [ ] add `facet_wrap(~Sex, ncol = 3)`
    - [ ] add `facet_grid(Sex~.)`
    - [ ] add `facet_grid(.~Sex)`
    - [ ] add `facet(.by = c(Sex))`


(c) Which of the following arguments to `geom_histogram()` could be present in the code that returned plot F? Choose all that apply.

    - [ ] `bins=12`
    - [ ] `bins=11`
    - [ ] `binwidth=1`
    - [ ] `binwidth=3`
    - [ ] `stat="identity"`

(d) Which of the following layers are required to produce plot F? Note that the plot uses the complete theme `theme_minimal()` and the font is 20pt Palatino. Choose all that apply.

    - [ ] `theme_minimal(use=TRUE)`
    - [ ] `theme_minimal(base_size = 20, base_family = "Palatino")`
    - [ ] `labs(title = "Histogram of speech rate")`
    - [ ] `font(size=20, family="Palatino")`
    - [ ] `scale_fill_manual(values = c("gray"))`

\clearpage

## 6 Data wrangling 

Section 6 makes use of the `durationsGe` dataset in the appendix.


(a) The `Sex` variable in the `durationsGe` dataset has the following distinct values: `"male" "female" NA`. How many rows would be in the object returned by the following code block? Write your answer in the box below.

    ```{r}
    #| echo: false

    library(languageR)
    ```

    ```r
    durationsGe %>% 
        filter(Sex %in% c("female")) %>%
        summarise(minBirthYear=min(YearOfBirth, na.rm=TRUE), .by=c(Sex))
    ```

    \answerbox


(b) True or false, the following code options are equivalent.

    ```r
    # option 1
    durationsGe %>% 
        select(Frequency) %>% 
        filter(Frequency > 40) %>%
        distinct()

    # option 2
    just_freq <- select(durationsGe, Frequency)
    freq_under_40 <- filter(just_freq, Frequency > 40)
    distinct(freq_under_40)
    ```

    - [ ] True
    - [ ] False

(c) Fill in the blank in the code below such that it returns a new column called "count", which counts of the number of rows in the `durationsGe` dataset per `Sex`

    ```r
    ratings %>% group_by(Sex) %>% summarise(_________)
    ```

    \answerbox

(d) True or false, the following code options are equivalent. 

    ```r
    # option 1
    durationsGe %>% 
        select(Freq=Frequency, Speaker:DurationOfPrefix) %>% 
        mutate(AgeInYears = 2023 - YearOfBirth, .before = Freq)

    # option 2
    durationsGe %>% 
        select(Frequency:DurationOfPrefix) %>% 
        rename(Freq = Frequency) %>%
        mutate(AgeInYears = 2023 - YearOfBirth, .before = 1)

    ```

    - [ ] True
    - [ ] False

\clearpage

*This page is intentionally blank in case you need scratch paper*

\clearpage

## Appendix A: Data


Sections 4-6 make use of `durationsGe` data in the `languageR` package. The dataset includes the duration of the prefix ge- in Dutch by various speakers from the Spoken Dutch Corpus. 

```{r}
library(languageR)
glimpse(durationsGe)
```

\clearpage 

## Appendix B: Plots section 4

```{r}
#| echo: false
#| warning: false
#| layout-ncol: 2
#| layout-valign: bottom
#| cap-location: bottom

ggplot(durationsGe, aes(x = DurationOfPrefix, fill = Sex)) +
    geom_density(fill = "lightgray") +
    theme_classic(base_size = 15) +
    labs(tag = "A", y = "")

ggplot(durationsGe, aes(x = DurationOfPrefix, fill = Sex)) +
    geom_density() +
    theme_classic(base_size = 15) +
    labs(tag = "B", y = "") +
    scale_fill_manual(values = c("white", "gray", "black"))
```



```{r}
#| echo: false
#| warning: false
#| layout-ncol: 2
#| layout-valign: bottom
#| cap-location: bottom
ggplot(durationsGe, aes(x = DurationOfPrefix)) +
    geom_rug() + 
    geom_histogram() +
    theme_classic(base_size = 12) +
    labs(tag = "C")
```

\clearpage

## Appendix C: Plots section 5

```{r}
#| echo: false
#| warning: false
#| layout-ncol: 2
#| layout-valign: bottom
#| cap-location: bottom

ggplot(durationsGe, aes(x = SpeechRate, y = DurationOfPrefix)) +
    geom_point(alpha = 0.5) + 
    geom_smooth(method = "lm") +
    theme_classic(base_size = 15) +
    labs(tag = "D")

ggplot(durationsGe, aes(x = SpeechRate, y = DurationOfPrefix)) +
    geom_point(alpha = 0.5) + 
    geom_smooth(method = "lm") + 
    facet_grid(. ~ Sex) +
    theme_classic(base_size = 15) +
    labs(tag = "E")

```

```{r}
#| echo: false
#| warning: false
#| layout-ncol: 2
#| layout-valign: bottom
#| cap-location: bottom
#| 
ggplot(data = durationsGe, mapping = aes(x = SpeechRate)) +
    geom_histogram(fill = "lightgray", bins = 12) +
    theme_minimal(base_family = "Palatino", base_size = 20) +
    labs(title = "Histogram of speech rate", tag="F") 
    
```


